The Nuclear Norm and Beyond: Dual Norms and Combinations for Matrix Optimization

### 2025-08-19 conf_article/neon.tex
Added proofs

We name the derived algorithm \emph{Neon}.

{\bf $\normfstar{\mM^{k}}^\dagger$ and F2-Muon}
We define $\normfstar{\cdot}$ as a convex combination of $\normn{\cdot}$ and $\normf{\cdot}$:

@@

We name the derived algorithm \emph{F2-Muon}. It turns out that F2-Muon is a convex combination of Normalized SGD and Muon, which is curious. The implications are significant and discussed in the following sections.

{\bf $\normftwo{\mM^{k}}^\dagger$ and F2-Neon}
We define $\normftwo{\cdot}$ as a convex combination of $\norms{\cdot}$ and $\normf{\cdot}$:

@@

where $\alpha \in [0, 1]$ defines a specific norm of $F2$-family.

@@

We name the derived algorithm \emph{F2-Neon}. It turns out that F2-Neon is a convex combination of Normalized SGD and Neon, which is curious. The implications are significant and discussed in the following sections.

@@

\subsection{Nuclear Neon}
Problem and update
\subsection{F*-Neon}
Motivation: we need updates of higher rank. Hence, F* (why? regularization?)
\subsection{F2-Muon}
Motivation: if the spectral norm (Muon) is dual to the nuclear norm (Neon), then what is dual to F*? F2-Muon! Derivation.

@@

    \section{Conclusion}
\begin{itemize}
    \item Future work: (L0, L1)-smoothness, probably extrapolation (to make dependency on eps milder, but we don't to spend memory here so never mind), hyperparameter-free (Bernstein2023)
\end{itemize}
\section{Appendix}
\subsection{Proofs}
\subsection{Norm derivations}
\begin{itemize}
    \item Dual to F* norm
    \item Bounds on rho for all 4 norms (with their tighness if possible)
\end{itemize}
\subsection{Updates derivations}
%yep, many KKT and so on
\begin{itemize}
    \item Neon
    \item F*-Neon
    \item F2-Muon
    \item F* choice from KKT
\end{itemize}
\subsection{}
~ Note: nuSAM and their update. They had no Lanczos
Technical details on experiments

>>>>

        We name the derived algorithm \emph{Neon}. In the section Matrix side of updates, we will discuss how to compute an update efficiently.

    {\bf $\normfstar{\mM^{k}}^\dagger$ and F-Muon}
        We define $\normfstar{\cdot}$ as a convex combination of $\normn{\cdot}$ and $\normf{\cdot}$:

@@

        We name the derived algorithm \emph{F-Muon}. It turns out that F-Muon is a convex combination of Normalized SGD and Muon, which is curious. The implications are significant and discussed in the following sections.

    {\bf $\normftwo{\mM^{k}}^\dagger$ and F-Neon}
        We define $\normftwo{\cdot}$ as a convex combination of $\norms{\cdot}$ and $\normf{\cdot}$:

@@

        where $\alpha \in [0, 1]$ defines a specific norm of F2-family.

@@

        We name the derived algorithm \emph{F-Neon}. It turns out that F-Neon is a convex combination of Normalized SGD and Neon, which is curious. The implications are significant and discussed in the following sections.

@@

\subsection{F*-Neon}
Motivation: we need updates of higher rank. Hence, F* (why? regularization?)
%\subsection{F-Muon}
%Motivation: if the spectral norm (Muon) is dual to the nuclear norm (Neon), then what is dual to F*? F-Muon! Derivation.

@@

\section{Conclusion}
\begin{itemize}
    \item Future work: (L0, L1)-smoothness, probably extrapolation (to make dependency on eps milder, but we don't to spend memory here so never mind), hyperparameter-free (Bernstein2023)
\end{itemize}
\section{Appendix}
\subsection{Norm derivations}
% \begin{itemize}
%    \item Dual to F* norm
%    \item Bounds on rho for all 4 norms (with their tighness if possible)
% \end{itemize}
\subsection{Updates derivations}

Derivation of \cref{eq:muon_update} follows from \cref{eq:nsgd_muon_update} with $\alpha=1$. Indeed, $\norms{\cdot}^\dagger=1\cdot\normn{\cdot} + 0 \cdot \normf{\cdot} $.

Derivation of \cref{eq:nsgd_muon_update}:
Since $\norm{\cdot}^\dagger = \normfstar{\cdot}^{\dagger\dagger} = \normfstar{\cdot}$, the goal is to reach $\normfstar{\mM^k} = \alpha \tr{\mSigma} + (1-\alpha) \normf{\mM^k}$.

Let us note that $\Delta = \alpha (\mU \mV^\top) + (1-\alpha) \frac{\mM^k}{\normf{\mM^k}}$ delivers this value. Indeed, by the trace property, $\<\mM^k, \Delta> =\<\mU \mSigma \mV^\top, \alpha \mU \mV^\top + (1-\alpha) \frac{\mU \mSigma \mV^\top}{\normf{\mM^k}}> = \alpha \tr \mSigma + (1-\alpha) \normf{\mM^k} = \normfstar{\mM^k}$, which completes the proof.

Derivation of \cref{eq:neon_update} follows from \cref{eq:nsgd_neon_update} with $\alpha=1$. Indeed, $\normn{\cdot}^\dagger=1\cdot\norms{\cdot} + 0 \cdot \normf{\cdot} $.

Derivation of \cref{eq:nsgd_neon_update}:
Since $\norm{\cdot}^\dagger = \normftwo{\cdot}^{\dagger\dagger} = \normftwo{\cdot}$, the goal is to reach $\normftwo{\mM^k} = \alpha \sigma_1 + (1-\alpha) \normf{\mM^k}$.

Let us note that $\Delta = \alpha (u_1 v_1^\top) + (1-\alpha) \frac{\mM^k}{\normf{\mM^k}}$ delivers this value. Indeed, by the trace property and singular vectors orthogonality, $\<\mM^k, \Delta> =\<\mU \mSigma \mV^\top, \alpha u_1 v_1^\top + (1-\alpha) \frac{\mU \mSigma \mV^\top}{\normf{\mM^k}}> = \alpha \tr \diag({\sigma_1, 0, \dots, 0}), + (1-\alpha) \normf{\mM^k} = \normftwo{\mM^k}$, which completes the proof.

\subsection{}
~ Note: nuSAM and their update. They had no Lanczos
Technical details on experiments
